---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# WhatToExpect

The goal of WhatToExpect is to take unaligned genomic sequencing reads and genrate
estimates for genome size, GC-content and (in the case of genomes with
isochore-like structures) the expected length-distributions for AT- and GC-rich
regions.

This is very much at an early stage of development, but feedback, PRs and issues
are welcome.

## Automated genome size estimates

I have an approximately 130kb 'fake' reference genome sitting the directory
`isnt/exdata`:


```{r example}
ref_len <- length(ape::read.dna("inst/exdata/fake_ref.fasta", "fasta"))
ref_len/1e3
```

From this, I have simulated 20 bp PE reads using wgsim. Again, the files are in
`inst/exdata`.

```{r reads}
reads <- list.files("inst/exdata", "*.fq", full.names=TRUE)
reads
```

So, let's use the reads to estimate the genome size. The first step is to use
`jellyfish` to count the occurrence of kmers. Because both the reads and the 
genome are are very small, we'll count 9-mers. 

```{r}
hist_file <- jfish_count_kmers(reads, k=9, starting_size="1M")
hist_file
```

This function returns the path to the binary file `jellyfish` stores the kmers
counts in. The function `read_jfish` lets us read that data into a 2-column
representation of the kmer spectrum. We can then use that dat to estimate the
genome size. As the name suggests, the function `est_genome_size` automates this
calculation. By setting `plot=TRUE` we can visualize the kmer-spectrum and see
the key parameters that are used for the estimate (so you can check everything
looks OK).

```{r}
kmer_hist <- read_jfish(hist_file)
size_est <- est_genome_size(kmer_hist, unit="b", plot=TRUE, x_cutoff=0.95)
size_est/1e3
```

In this case, we get a very nice estimate of the "true" genome size from the
sequencing reads:


```{r}
dotchart(c(size_est,ref_len), 
         labels=c("Estimate", "True genome size"),
         xlim=c(0,ref_len))
```
